Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-03-25 main.Rmd
Issue 11: tables need cleaning up

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
library(ggridges)
library(tsibble)
library(gravitas)
library(tsibble)
library(ggpubr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = "figure/", fig.align = "center", fig.show = "hold",
  cache = TRUE, cache.path = "cache/",
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
```

@@

<!-- some discussions on primary units of calendar -->
**Example:** So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar. But it is far from being the only one. <!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->The day is the basic unit of time underlying most calendars [@Reingold2001-kf]. Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years. The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds". Nevertheless, for any calendar a hierarchy table can be defined. For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun. \autoref{tab:tab-mayan} represents the the hierarchy table for the Mayan calendar.

```{r tab-mayan}
table_mayan <- tibble(G = c("kin", "uinal", "tun", "katun", "baktun"),
                      C = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1),
                      K = c(20, 18, 20, 20, 1)) 
kable(table_mayan, 
      format = "latex",
      booktabs = TRUE,
      caption = "Hierarchy table for Mayan calendar. $G$ and $C$ represent the linear and single order-up cyclic granularities respectively. All cyclic granularities are circular with $K$ as the period length.")
```

@@

```{r tab-gregorian}
table_greg <- tibble(G = c("minute", "hour", "day", "month", "year"),
                      C = c("minute-of-hour","hour-of-day","day-of-month", "month-of-year", 1),
                     K = c(20, 18,"k(day, month)", 20, 1))

kable(table_greg, 
      format = "latex",
      booktabs = TRUE,
      caption = "Illustrative hierarchy table for Gregorian calendar. $G$ and $C$ represent the linear and single order-up cyclic granularities respectively. All these cyclic granularities are circular with $K$ as the period length except for day-of-month which is quasi-circular.")
```

@@

<!-- # how data looks -->
Smart meters provide large quantities of measurements on energy usage for households across Australia. One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014. <!--It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.-->The idea here is to show how to look at the energy consumption across different cyclic granularities in a systematic way to identify different behavioral patterns.
<!-- and identify the extreme and regular households. -->

### Data structure for visualization:

The data structure for a sample of 50 households from that trial is depicted in \autoref{tab:smart-data} where the variable `reading_datetime`, `customer_id` and `general_supply_kwh` denote the index, key and measured variable of the tsibble.
<!-- We take a random sample of 50 households to see how we can achieve that.  -->

```{r smart-data}
load("data/sm_cust50.Rdata")

smart_data <- smart_meter <- sm_cust50 %>%
  as_tsibble() %>%
  dplyr::select(
    customer_id,
    reading_datetime,
    general_supply_kwh
  ) %>% head(n = 5)

  knitr::kable(smart_data,
               format = "latex",
               booktabs = TRUE,
               caption = "Smart meter data structure with electricity demand for 50 households from SGSC project for every 30 minutes from 2012 to 2014.")
```

@@

```{r search_gran_limit2, echo=FALSE, eval = FALSE}
smart_meter10 %>% search_gran(
  highest_unit = "month",
  filter_out = c("hhour", "fortnight"))%>%
  knitr::kable(format = "latex",
      booktabs = TRUE)
```

@@

  knitr::kable(smart_harmony,
               format = "latex",
               booktabs = TRUE,
                caption = "Harmonies with two cyclic granularity one placed on the facet and the other on x-axis with the corresponding levels. Out of 30 possible combinations of cyclic granularities that could be plotted only 13 are harmonies.")
```

@@

```{r hierarchy-cric}
library(gravitas)
library(tibble)
hierarchy_model <- tibble::tibble(
  G = c("over", "inning", "match", "season"),
  C = c("over-of-inning", "inning-of-match", "match-of-season", 1),
  K = c(20, 2, "k(match, season)", 1)
)
knitr::kable(hierarchy_model,
             format = "latex",
             booktabs = TRUE,
             caption = "Hierarchy table for cricket where each over is nested within innings, innings is nested within match and so on. Cyclic granularity over-of-inning, inning-of-match are assumed to be circular which means each inning will have 20 overs and each match has 2 innings and  match-of-season is assumed to be quasi-periodic. ")
```

>>>>

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
library(ggridges)
library(tsibble)
library(gravitas)
library(tsibble)
library(ggpubr)
library(xtable)
library(kableExtra)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = "figure/", fig.align = "center", fig.show = "hold",
  cache = TRUE, cache.path = "cache/",
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
```

@@

<!-- some discussions on primary units of calendar -->
**Example:** So far we have used example of cyclic granularities from Gregorian calendar as it is the most widely used calendar. But it is far from being the only one. <!--All calendars fall under three types - solar, lunar or lunisolar/solilunar-->The day is the basic unit of time underlying most calendars [@Reingold2001-kf]. Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years. The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds". Nevertheless, for any calendar a hierarchy table can be defined. For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured such that 1 kin = 1 day; 1 uinal = 20 kin; 1 tun  = 18 uinal; 1 katun = 20 tun and 1 baktun = 20 katun. \autoref{tab:tab-mayan} represents the hierarchy table for the Mayan calendar.

```{r tab-mayan}
table_mayan <- tibble(`linear (G)` = c("kin", "uinal", "tun", "katun", "baktun"),
                      `single-order-up cyclic (C)` = c("kin-of-uinal","uinal-of-tun"," tun-of-katun", "katun-of-baktun", 1),
                       `period length/conversion operator (K)`= c(20, 18, 20, 20, 1)) 
kable(table_mayan, 
      format = "latex",
      booktabs = TRUE,
      align = "c",
      caption = "Hierarchy table for Mayan calendar with all circular single-order-up granularities.")
  #row_spec(0, bold = TRUE) %>% 
  #row_spec(1:(nrow(table_mayan)-1), hline_after = T) %>% kable_styling(latex_options = "striped")
```

@@

```{r tab-gregorian}
table_greg <- tibble(`linear (G)` = c("minute", "hour", "day", "month", "year"),
                      `single-order-up cyclic (C)` = c("minute-of-hour","hour-of-day","day-of-month", "month-of-year", 1),
                     `period length/conversion operator (K)` = c(20, 18,"k(day, month)", 20, 1))

kable(table_greg, 
      format = "latex",
      booktabs = TRUE,
      caption = "Illustrative hierarchy table for Gregorian calendar with both circular and quasi-circular single-order-up granularities.") %>% 
  #row_spec(0, bold = TRUE) %>%
   kable_styling()
```

@@

<!-- # how data looks -->
Smart meters provide large quantities of measurements on energy usage for households across Australia. One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City project in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption for every half hour from February 2012 to March 2014. <!--It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units.-->The idea here is to show how to look at the energy consumption across different cyclic granularities in a systematic way to identify different behavioral patterns.

<!-- and identify the extreme and regular households. -->

<!-- ### Data structure for visualization: -->

<!-- The data structure for a sample of 50 households from that trial is depicted in \autoref{tab:smart-data} where the variable `reading_datetime`, `customer_id` and `general_supply_kwh` denote the index, key and measured variable of the tsibble. -->
<!-- <!-- We take a random sample of 50 households to see how we can achieve that.  -->

```{r smart-data}
# load("data/sm_cust50.Rdata")
# 
# smart_data <- smart_meter <- sm_cust50 %>%
#   as_tsibble() %>%
#   dplyr::select(
#     customer_id,
#     reading_datetime,
#     general_supply_kwh
#   ) %>% head(n = 5)
# 
#   knitr::kable(smart_data,
#                format = "latex",
#                booktabs = TRUE,
#                caption = "Smart meter data structure with electricity demand for 50 households from SGSC project for every 30 minutes from 2012 to 2014.")
```

@@

```{r search_gran_limit2, echo=FALSE, eval = FALSE}
smart_meter10 %>% search_gran(
  highest_unit = "month",
  filter_out = c("hhour", "fortnight"))%>%
  knitr::kable(format = "latex",
      booktabs = TRUE) %>%
  #row_spec(0, bold = TRUE)%>%
  kable_styling()
```

@@

  knitr::kable(smart_harmony,
               format = "latex",
               booktabs = TRUE,
                caption = "Harmonies with two cyclic granularity one placed on facet and the other on x-axis.") %>% 
    #row_spec(0, bold = TRUE) %>% 
    kable_styling() %>% 
     add_footnote("NOTE: Out of 30 possible combinations of cyclic granularities, only these 13 are harmony pairs.",
                  notation="none")
```

@@

```{r hierarchy-cric}
library(gravitas)
library(tibble)
hierarchy_model <- tibble::tibble(
  `linear (G)` = c("over", "inning", "match", "season"),
  `single-order-up cyclic (C)` = c("over-of-inning", "inning-of-match", "match-of-season", 1),
  `period length/conversion operator (K)` = c(20, 2, "k(match, season)", 1)
)
knitr::kable(hierarchy_model,
             format = "latex",
             booktabs = TRUE,
             caption = "Hierarchy table for cricket with each over is nested within innings, innings nested within match and match within season.") %>% 
  #row_spec(0, bold = TRUE)%>%
  kable_styling()
```
